---
title: "Take-home exercise 4"
description: |
  ISSS608 Visual Analytics Take-home Exercise 4
author:
  - name: Jiarui Cui
    url: www.linkedin.com/in/jiarui-cui-482232195
    affiliation: School of Computing and Information Systems
    affiliation_url: https://scis.smu.edu.sg/
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

This take-home exercise aims to reveal the daily routines of two selected participant of the city of Engagement, Ohio USA by using use ViSIElse and other appropriate visual analytics methods.

# Importing Packages

In this take-home exercise, the following R packages will be used:
 - sf, an R package specially designed to handle geospatial data in simple feature objects.
 
```{r}
packages = c('lubridate', 'tidyverse', 
             'readr','tmap','sf','sftime',
             'clock','rmarkdown','ViSiElse',
             'fastDummies','ggplot2')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

# Importing data

```{r eval = FALSE}
logs_fread <- list.files(path = "Data/ActivityLogs/",
                  pattern = "*.csv", 
                  full.names = T) %>% 
  map_df(~fread(.))
```
```{r}
buildings <- read_sf("Data/Buildings.csv", 
                   options = "GEOM_POSSIBLE_NAMES=location")
```

```{r eval = FALSE,echo = FALSE}
write_rds(logs_fread,"Data/logs_fread.rds")
```

```{r echo = FALSE}
logs_fread <- read_rds('Data/logs_fread.rds')
```

# Data preparation

Create label **date**, **time**, and **hours**.

```{r}
logs_fread$month <- factor(month(logs_fread$`timestamp`), 
                    levels=1:12, 
                    labels=month.abb, 
                    ordered=TRUE) 

logs_March <- logs_fread %>% filter(month == 'Mar')
logs_March$date <- as_date(logs_March$`timestamp`)
logs_March$time <- format(logs_March$timestamp, format = "%H:%M:%S")
logs_March$hour <-hour(logs_March$`timestamp`)
```

Change the time into numerical value - number of minutes

```{r}
log_participant4 = logs_March %>% mutate(date = date(timestamp),
         time = (hour(timestamp)*60+minute(timestamp))) %>%
  filter(participantId== 4)
log_participant16 = logs_March %>% mutate(date = date(timestamp),
         time = (hour(timestamp)*60+minute(timestamp))) %>%
  filter(participantId== 16)
```

Calculate their daily average balance.

```{r}
participant4_balance <- log_participant4%>% 
  group_by(date) %>% 
  summarize(Balance = mean(availableBalance))

participant16_balance <- log_participant16%>% 
  group_by(date) %>% 
  summarize(Balance = mean(availableBalance))
```

## Balance comparison

The line chart below presents the available balance for participant 4 and 16 throughout March 2022. 

```{r}
balance_comparison <- merge(participant4_balance,participant16_balance,
                            by="date")
ggplot(balance_comparison, aes(date)) + 
  geom_line(aes(y = Balance.x, colour = "participant 4")) + 
  geom_line(aes(y = Balance.y, colour = "participant 16"))
  
```

# Daily routine comparison

```{r}
grouped4 <- log_participant4 %>% 
  count(hour,currentMode) %>% 
  ungroup() %>%
  na.omit()
ggplot(grouped4, 
       aes(hour,currentMode, 
           fill = n)) + 
geom_tile(color = "white", 
          size = 0.1) + 
coord_equal() +
scale_fill_gradient(name = "Frequencies",
                    low = "sky blue", 
                    high = "dark blue") +
labs(x = 'Hour', 
     y = NULL, 
     title = "Monthly routine for participant 4") +
theme(axis.ticks = element_blank(),
      plot.title = element_text(hjust = 0.5),
      legend.title = element_text(size = 4),
      legend.text = element_text(size = 3) )

grouped16 <- log_participant16 %>% 
  count(hour,currentMode) %>% 
  ungroup() %>%
  na.omit()
ggplot(grouped16, 
       aes(hour,currentMode, 
           fill = n)) + 
geom_tile(color = "white", 
          size = 0.1) + 
coord_equal() +
scale_fill_gradient(name = "Frequencies",
                    low = "sky blue", 
                    high = "dark blue") +
labs(x = 'Hour', 
     y = NULL, 
     title = "Monthly routine for participant 16") +
theme(axis.ticks = element_blank(),
      plot.title = element_text(hjust = 0.5),
      legend.title = element_text(size = 4),
      legend.text = element_text(size = 3) )
```

```{r}
get_home_time4 <- (log_participant4%>%
          filter(currentMode == 'AtHome'))[, c('date','time')]%>%
  group_by(date) %>% 
  summarize(get_home = min(time))

leave_home_time4 <- (log_participant4%>%
          filter(currentMode == 'AtHome'))[, c('date','time')]%>%
  group_by(date) %>% 
  summarize(leave_home = max(time))

start_travel_time4 <- (log_participant4%>%
          filter(currentMode == 'Transport'))[, c('date','time')]%>%
  group_by(date) %>% 
  summarize(start_travel = min(time))

end_travel_time4 <-(log_participant4%>%
          filter(currentMode == 'Transport'))[, c('date','time')]%>%
  group_by(date) %>% 
  summarize(end_travel = max(time))

start_work_time4 <- (log_participant4%>%
          filter(currentMode == 'AtWork'))[, c('date','time')]%>%
  group_by(date) %>% 
  summarize(start_work = min(time))

end_work_time4 <-  (log_participant4%>%
          filter(currentMode == 'AtWork'))[, c('date','time')]%>%
  group_by(date) %>% 
  summarize(end_work  = max(time))

get_restaurant_time4 <-  (log_participant4%>%
          filter(currentMode == 'AtRestaurant'))[, c('date','time')]%>%
  group_by(date) %>% 
  summarize(get_restaurant = min(time))

leave_restaurant_time4 <- (log_participant4%>%
          filter(currentMode == 'AtRestaurant'))[, c('date','time')]%>%
  group_by(date) %>% 
  summarize(leave_restaurant  = max(time))

start_recreation_time4 <- (log_participant4%>%
          filter(currentMode == 'AtRecreation'))[, c('date','time')]%>%
  group_by(date) %>% 
  summarize(start_recreation = min(time))

end_recreation_time4 <- (log_participant4%>%
          filter(currentMode == 'AtRecreation'))[, c('date','time')]%>%
  group_by(date) %>% 
  summarize(end_recreation  = max(time))

get_home_time16 <- (log_participant16%>%
          filter(currentMode == 'AtHome'))[, c('date','time')]%>%
  group_by(date) %>% 
  summarize(get_home = min(time))

leave_home_time16 <- (log_participant16%>%
          filter(currentMode == 'AtHome'))[, c('date','time')]%>%
  group_by(date) %>% 
  summarize(leave_home = max(time))

start_travel_time16 <- (log_participant16%>%
          filter(currentMode == 'Transport'))[, c('date','time')]%>%
  group_by(date) %>% 
  summarize(start_travel = min(time))

end_travel_time16 <-(log_participant16%>%
          filter(currentMode == 'Transport'))[, c('date','time')]%>%
  group_by(date) %>% 
  summarize(end_travel = max(time))

start_work_time16 <- (log_participant16%>%
          filter(currentMode == 'AtWork'))[, c('date','time')]%>%
  group_by(date) %>% 
  summarize(start_work = min(time))

end_work_time16 <-  (log_participant16%>%
          filter(currentMode == 'AtWork'))[, c('date','time')]%>%
  group_by(date) %>% 
  summarize(end_work  = max(time))

get_restaurant_time16 <-  (log_participant16%>%
          filter(currentMode == 'AtRestaurant'))[, c('date','time')]%>%
  group_by(date) %>% 
  summarize(get_restaurant = min(time))

leave_restaurant_time16 <- (log_participant16%>%
          filter(currentMode == 'AtRestaurant'))[, c('date','time')]%>%
  group_by(date) %>% 
  summarize(leave_restaurant  = max(time))

start_recreation_time16 <- (log_participant16%>%
          filter(currentMode == 'AtRecreation'))[, c('date','time')]%>%
  group_by(date) %>% 
  summarize(start_recreation = min(time))

end_recreation_time16 <- (log_participant16%>%
          filter(currentMode == 'AtRecreation'))[, c('date','time')]%>%
  group_by(date) %>% 
  summarize(end_recreation  = max(time))
```

```{r}
participant4_list <- list(get_home_time4, leave_home_time4, 
              start_travel_time4,end_travel_time4,
              start_work_time4,end_work_time4,
              get_restaurant_time4,leave_restaurant_time4,
              start_recreation_time4,end_recreation_time4)
participant4_routine <- participant4_list %>% reduce(full_join, by='date')

participant16_list <- list(get_home_time16, leave_home_time16, 
              start_travel_time16,end_travel_time16,
              start_work_time16,end_work_time16,
              get_restaurant_time16,leave_restaurant_time16,
              start_recreation_time16,end_recreation_time16)
participant16_routine <- participant16_list %>% reduce(full_join, by='date')
```

```{r}
visielse(participant4_routine)
visielse(participant16_routine)
```